---
title: "HTML EDA:<br>Fast Food Nutrition"
author: "Ahmir Postell"
format: 
  html:
    code-overflow: wrap
    toc: true
    toc-depth: 4
    toc_float: true
    toc-title: Table of Contents
editor: 
  markdown: 
    wrap: 72
---

# 1. The Research Problem

My project addresses a significant concern within the student community
at William and Mary regarding the transition from mandatory meal plans
to optional ones starting in the junior year. This shift often leads to
increased reliance on fast food options among students, raising
questions about the nutritional and financial implications compared to
the university meal plan. My project aims to provide clarity on these
matters by conducting a comprehensive analysis of the nutritional and
financial disparities between meal plan offerings and commonly chosen
fast food items. Additionally, the project will delve into the
variations among fast food establishments, shedding light on the diverse
choices available and their respective impacts. By undertaking this
investigation, the project seeks to equip students with tangible
insights into the decision-making process surrounding meal plans versus
fast food, ultimately facilitating more informed choices regarding their
dietary and financial well-being during their academic journey.

# 2. Research Questions

The primary objectives of this study are to glean insights into the
nutritional offerings of fast food establishments in proximity to
William and Mary's campus and to contrast these offerings with those
available through a meal plan. Additionally, the study aims to determine
the financial implications associated with students opting out of a meal
plan in favor of fast food consumption. The overarching research
questions guiding this investigation are as follows:

Which fast food restaurant near William and Mary’s campus offers the
most nutritious and well balanced food? What is the frequency at which a
student can consume fast food and still maintain financial advantages
compared to having a meal plan?

These research questions are designed to provide quantitative
understanding of the nutritional content and financial considerations
associated with fast food consumption relative to meal plan options,
thereby informing students' decision-making processes regarding their
dietary and financial well-being during their academic tenure.

# 3. Data Source

The original source of my data is the TidyTuesday GitHub page, which
compiled information sourced from fastfoodnutrition.org. I have
personally verified the reliability of the source by cross-referencing
some of the data. Additionally, fastfoodnutrition.org is an established
platform known for its comprehensive and accurate nutritional
information on various fast food items. Moreover, the use of an official
".org" domain adds to the credibility of the source.

However, upon closer inspection, this data set was not adequate to
address my research question. I decided to scrape fastfoodnutrition.org
personally to ensure I was able to obtain all the information I needed.

To enhance my current dataset, I will clean my data set of unnecessary
data and I intend to incorporate additional data regarding the price of
individual items, as well as nutritional and financial data specific to
William and Mary's dining options. For this purpose, I will utilize
information available on William and Mary's official website,
considering it the most reliable and accurate source for such
institutional data.

# 4. Exploratory Data Analysis

```{r setup}
#Load packages and set any options or constants, ensuring all chunks of code are rendered; be sure to comment your code throughout the file
library(dplyr)
library(tidyverse)
```

```{r}
# reading in data set
data <- read.csv("~/Desktop/Spring 2024/Data Analytics with R/Final Project/FastFood_v_campus_meal_plan/Data/data.csv")
```

## 4a. Overview

## 4b. Notable Variables

### Notable Variable 1: Daily Percent of Calories

Fast food restaurants are known for provide a large amount of calories
and this can be a significant issue to many people, including college
students. To make interpreting my data set easier, I created a variable
to represent the amount of a healthy daily calorie intake is satisfied
by each item. Through research, I learned that, on average, women should
intake at least 2,000 calories per day and men should have at least
2,500 calories in order to maintain weight. With this information, I
created a variable by dividing the calories of each item by 2,250 (the
middle) and multiplying by 100 to get the percent of the a healthy
amount of calories each item takes up. I believe this knowledge aides in
interpreting if item are "too large" to consume frequently.

```{r}
#Process for creating or modifying a variable...
data$Daily_Percent_Calories <- round((data$Calories / 2250) * 100, digits = 0)

```

### Notable Variable 2: Percent Trans Fat

There are multiple types of fat and trans fat is regarded as the worst
fat. To further visualize the harm an item might have, I created a
variable by finding the ratio of Fats that is made up of trans fats,
allowing me to further analyse which items are healthy and which aren't.

```{r}
#Process for creating or modifying a variable...
data$Percent_Bad_Fat <- round((data$Trans_Fat.g / data$Total_Fat.g) * 100, digits = 0)

```

### Notable Variable 3: High Colesterol

I created this variable to determine if an item is high in cholesterol.
For individuals with high cholesterol, I felt this information necessary
in making a choice between having a meal plan and not having one. In
order to create this variable, I first researched and found that one
should limit themselves to 300mg of cholesterol a day to avoid health
complications. Considering one item as a meal and assuming a 3-meal
diet, I determined that an item containing 90mg of cholesterol is
considered "high in cholesterol". I use 90mg instead of 100mg to adjust
for those already at risk for heart disease.

```{r}

data <- data %>%
  mutate(High_Cholesterol = ifelse(Cholesterol.mg > 90, "yes", "no"))


```

## 4c. Summary Statistics

***Please limit the use of `summary()` and `describe()` to one variable
each. Other options include calculating correlations, calculating
frequency counts, or identifying outliers (`boxplot.stats()`).***

### Variable 1: Cholesterol per Restaurant

This variable calculates the average amount of cholesterol per
restaurant.

```{r}
# Calculate summary statistics and comment your code so I can follow your process
cholesterol_per_restaurant <- data %>%
  group_by(Restaurant) %>%
  summarise(avg_cholesterol = round(mean(Cholesterol.mg, na.rm = TRUE), digits = 0)) %>%
  ungroup()

# View the result
cholesterol_per_restaurant
```

### Variable 2: Calorie to Fat per Restaurant

This variable calculates the average number of calories and the average
amount of calories from fat per restaurant. I use these numbers to by
the mean ration of calories made up by fat. In retrospect, I believe I
could have done this in an easier way.

```{r}
# Calculate summary statistics and comment your code so I can follow your process
calorie_to_fat_per_restaurant <- data %>%
  group_by(Restaurant) %>%
  summarise(
    # calculate average number of calorie and round to 2 decimal places
    avg_num_calories = round((mean(Calories, na.rm = TRUE)), digits = 2),
    # calculate the average number of calories from fat and round to 2 decimal places
    calories_from_fat = round((mean(Calories_From_Fat, na.rm = TRUE)), digits = 2),
    # find the ration of calories that come from fat and round to 0 decimal place b/c it is a percentage
    calories_to_fat_ratio = round((calories_from_fat / avg_num_calories) * 100, digits = 0)) %>%
  
  # remove the first 2 columns created during calculations
  select(-avg_num_calories) %>%
  select(-calories_from_fat) %>%
  ungroup()

calorie_to_fat_per_restaurant
```

### Variable 3: Items per Restaurant

I created this variable to visualize the variety of options available
for each restaurant. It counts the number of items within my data set
grouping by restaurant to get the total amount of items associated with
that restaurant.

```{r}
# Calculate summary statistics and comment your code so I can follow your process
num_items_per_restaurant <- data %>%
  group_by(Restaurant) %>%
  summarise(item_count = n()) %>%
  ungroup()

# View the result
num_items_per_restaurant
```

### Variable 4: Cholesterol by Type

This variable calculates the average amount of cholesterol in regards to
type of food that each item fall into using the "Type" column in my data
set. I create this to fully understand which types of fast food
correlate to more cholesterol. This will also help with visualizations
regarding what types of fast food to avoid in order to cut back on
cholesterol.

```{r}
# Calculate summary statistics and comment your code so I can follow your process
cholesterol_by_type <- data %>%
  group_by(Type) %>%
  summarise(cholesterol = round(mean(replace(Cholesterol.mg, is.na(Cholesterol.mg), 0)), digits = 0)
            ) %>%
  ungroup()

cholesterol_by_type
```

### Variable 5: Protein by Restaurant

This variable calculates the mean amount of protein available per
restaurant. For those wanting to build muscle, I felt it necessary to
determine with restaurants offer in terms of protein as that is an
important nutrient for muscle recovery.

```{r}
# Calculate summary statistics and comment your code so I can follow your process
protein_by_restaurant <- data %>%
  group_by(Restaurant) %>%
  summarise(avg_protein = round(mean(Protein.g, na.rm = TRUE), digits = 0)
            ) %>%
  ungroup()

protein_by_restaurant
```

5-10 sentences about any notable or surprising findings, as well as
potential outliers...

Although due to my ignorance, I was surprised to find that bread doesn't
contain cholesterol. I was not knowledgeable of where each nutrient is
found in food and found it notable that I am just finding this out now.
I also was surprised to learn that subway had the highest average amount
of protein. I thought that popeyes would be the highest as they nearly
serve only chicken. I also thought that chipotle would be higher as I
know many gym goers opt for chiptole as their post workout meal.

## 4d. Data Visualizations

### Bar Plot 1

This plot represents the total number of items offered per restaurant.
This is pretty straight forward however, I think it is notable that
starbucks offers the largest amount of item. I imagine this is because I
recorded data for each size per drink available.

```{r}
#Creating the visualization...
ggplot(data = num_items_per_restaurant, aes(x = Restaurant, y = item_count)) +
  geom_col(fill = "darkred") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "Number of Items Available",title = "Amount of Items per Restaurant")

```

### Bar Plot 2

This plot visualizes the ratio of amount of calories derived from fat.
This results of this visualization are expect. Fast food as a whole has
a large amount of calories derived from fat. It is also unsurprising to
find that Five Guys has the largest percent of total calories derived
from fat.

```{r}
#Creating the visualization...
ggplot(data = calorie_to_fat_per_restaurant, aes(x = Restaurant, y = calories_to_fat_ratio)) +
  geom_col(fill = "darkgreen") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "Percent of Calories From Fat",title = "Percent Calories From Fat per Restaurant")

```

### Histogram 1

This histogram displays the frequency of items that reach a certain
percent of a healthy daily calorie intake. I found the results fairly
surprising as I expected the plot to be more skewed to the 50% range. It
seem that the majority of items contain below the 25% of daily calorie
intake.

```{r}
#Creating the visualization...
plot_data <- data %>%
  filter(!is.na(Daily_Percent_Calories))

ggplot(data = plot_data, aes(x = Daily_Percent_Calories)) +
  geom_histogram(boundary = 0, bins = 8, fill = "blue", color = "black") + 
  scale_x_continuous(breaks = seq(0, 100, by = 25)) +
  labs(y = "Frequency", x = "% of Daily Calorie Intake", title = "Frequency of Daily Calorie Intake")
```

### Histogram 2

This plot visualizes the frequency of ranges of total carbohydrates
among my data. Most fast food items are below 100 mg of carbohydrates,
however I believe this is a high amount in general. I feel I need to
look closer at the the higher frequencys to conclude if these foods are
high in carbohydrates or not.

```{r}
#Creating the visualization...
plot_data <- data %>%
  filter(!is.na(Total_Carbohydrates.g))

ggplot(data = plot_data, aes(x = Total_Carbohydrates.g)) +
  geom_histogram(boundary = 0, bins = 12, fill = "purple", color = "gray") +
  labs(y = "Frequency", x = "Total Carbohydrates", title = "Frequency of Total Carbohydrates")

```

### Box Plot 1

This plot shows the distribution of sugars among my data. I thing this
shows the lack of cleaning done to my data. There are a lot of outlier
and I feel like I can't intepret much from this data, outside of meaning
my data set include a lot of extra data.

```{r}
#Creating the visualization...
plot_data <- data %>%
  filter(!is.na(Sugars.g))

ggplot(plot_data, aes(x = Restaurant, y = Sugars.g)) +
  geom_boxplot(fill = "violet") +
  scale_x_discrete(limits = rev) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "Sugars (g)", title = "Distribution of Sugars")
```

### Box Plot 2

This plot is visualizes the distribution of calories from fat per
restuarant with my data set. Again, I have the same problem interpreting
this plot and will need to clean my data to get more information from
it.

```{r}
#Creating the visualization...
plot_data <- data %>%
  filter(!is.na(Calories_From_Fat))

ggplot(plot_data, aes(x = Restaurant, y = Calories_From_Fat)) +
  geom_boxplot(fill = "red") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "Calories From Fat", title = "Distribution of Calories From Fat")

```

### Scatter Plot 1

This scatter plot displays the correlation of calories and protein
throughout my data set. Using it, I can determine that there is a
positive correlation. This means that I can expect higher calorie fast
food items to contain more protein to an extent.

```{r}
#Creating the visualization...
plot_data <- data %>%
  filter(!is.na(Calories)) %>%
  filter(!is.na(Protein.g))

ggplot(plot_data, aes(x = Calories, y = Protein.g)) +
  geom_point(shape = 21, color = "gray", fill = "gold") +
  geom_smooth(method = "lm", formula = y~x, se = FALSE,
              color = "red", linetype = "dashed", linewidth = .5) +
  labs(y = "Protein (g)", title = "Calories v. Protein in Fast Food")
```

### Scatter Plot 2

This scatter plot represent the correlation of total fats and trans fats
within my data set. I find this plot harder to interpret as it doesn't
quite seem like they correlate.

```{r}
#Creating the visualization...
plot_data <- data %>%
  filter(!is.na(Total_Fat.g)) %>%
  filter(!is.na(Trans_Fat.g))

ggplot(plot_data, aes(x = Total_Fat.g, y = Trans_Fat.g)) +
  geom_point(shape = 21, color = "blue", fill = "white") +
  geom_smooth(method = "lm", formula = y~x, se = FALSE,
              color = "red", linetype = "dashed") +
  labs(y = "Trans Fat (g)", x = "Total Fat (g)", title = "Total Fat v. Trans Fat in Fast Food")
```

### Wild Card Plot 1

This box plot represents the distribution of cholesterol among types of
food represented by my data set. It seems that foods in the chicken
category are pretty well distributed. Its also interesting that
breakfast has the most outlier. It's worth noting that I limited this
display to the items with the most cholesterol to ensure readability.

```{r}
#Creating the visualization...
plot_data <- data %>%
  top_n(200, Cholesterol.mg)

ggplot(plot_data, aes(x = Type, y = Cholesterol.mg)) +
  geom_boxplot(fill = "gray", color = "darkblue") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "Cholestorol", title = "Distribution of Cholestorol by Type")
```

### Wild Card Plot 2

This bar plot shows the average amount of protein per restaurant. I used
this to visualize my summarized data and was surprised to see that the
difference between the highest and lowest average protein wasn't as
large as I had thought.

```{r}
#Creating the visualization...
ggplot(data = protein_by_restaurant, aes(x = Restaurant, y = avg_protein)) +
  geom_col(fill = "darkorange", color = "black") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "Mean Protein",title = "Average Protein per Item")
```

## 4e. Challenges

**Challenge 1:** The biggest challenge I've faced so far is
consistencies when finding the data. My original data set did not
consist of all the fast food restaurant that I wanted to look at and
although I could find alternative data sets for the individual
restaurants, the columns and data included in the alternative data sets
would be very different in terms of actual data given and accuracy of
data. Because of this, I needed to use web scraping to extract the data
myself. This in itself was very challenging as I had to learn how to web
scrape using r before obtaining and analyzing the data. After, scraping
all the nutritional data, I realize that the library I was using (rvest)
would not work for scraping a the website I intended to use for scraping
the prices of each item because it renders using javascript. I spent a
lot of time trying to work through this problem and have not yet found a
solution.

**Challenge 2:** Challenge 1 led to another challenge: time. I spent a
large amount of time learning and scraping the data to build my data
set. Because of this, my time was very limited in analyzing and
interpreting the actual data. I also didn't have time to clean and
finalize my data set. Consequently, I had to adjust my time management
and priorities to continue making progress throughout this project.

# 5. Preliminary Conclusions and Next Steps

So far, this most notable thing that I learned from my data is that it
needs to be cleaned. I noticed that when scraping my data, I created
entries for each toppings and ingredients.I believe that this created a
lot of skewed results in my data. Despite this, I believe that my some
of my finding won't change after cleaning my data set such as the
restaurant that offer the largest amount of protein and calories. I
found most of my findings to be in-line with what I thought common fast
food offered; however, I think it is interesting how much cholesterol
some items contain.

For my next steps, I'd like to clean my data, add additional data regard
price, and pull data regarding options offered by William and Mary's
meal plan. I believe this is a good start, but I have a lot of work
before I can draw conclusions that will answer my research questions.
Specifically, I want to analyze how price comes into play throughout my
data. I also hope to have more unique visualizations using scatter plots
and am particularly interested in exploring pie charts if possible.
